---
title: "QC Report"
author: "r-scrna development team"
date: "`r Sys.Date()`"
output:
  openintro::lab_report
---

# 1. Executive Summary

## Overview

This report summarizes the quality control (QC) metrics and filtering results for the single-cell RNA-seq dataset.

## Analysis Information

- **Project**: `r params$project$name`
- **Analysis Date**: `r format(Sys.Date(), "%Y-%m-%d")`

## Summary Statistics

| Metric | Value |
|--------|-------|
| **Total Cells (Initial)** | `r qc_stats[1, 3]` |
| **Total Genes (Initial)** | `r qc_stats[2, 4]` |
| **Mean Features/Cell** | `r qc_stats[3, 3]` |
| **Median Features/Cell** | `r qc_stats[4, 3]` |
| **Mean Counts/Cell** | `r qc_stats[5, 3]` |
| **Median Counts/Cell** | `r qc_stats[6, 3]` |
| **Mean MT Percentage** | `r qc_stats[7, 3]` |
| **Median MT Percentage** | `r qc_stats[8, 3]` |

# 2. Filtering Criteria

Applied filtering thresholds based on configuration settings:

| Parameter | Threshold | Description |
|-----------|----------|-------------|
| **Min Features/Cell** | `r config$qc$min_features_per_cell` | Minimum genes per cell |
| **Max Features/Cell** | `r ifelse(is.null(config$qc$max_features_per_cell), "Not set", config$qc$max_features_per_cell)` | Maximum genes per cell (optional) |
| **Max MT Percentage** | `r config$qc$max_mt_pct` | Maximum mitochondrial percentage |

### Cells Filtered

- **Total Cells (After Filtering)**: `r qc_stats[1, 3]`
- **Cells Removed**: `r qc_stats[1, 3] - ncol(obj_after)`

```{r, eval=FALSE}
library(Seurat)
# Assuming obj_after is the filtered object
cat("Cells after filtering:", ncol(obj_after), "\n")
```

# 3. QC Visualizations

## 3.1 Violin Plots

### 3.1.1 Features/Cell Distribution

```{r, eval=FALSE, fig.width=10, fig.height=6}
library(Seurat)
VlnPlot(obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
        ncol = 3, pt.size = 0.1) +
  plot_annotation(title = "QC Metrics Violin Plots")
```

### 3.1.2 MT Percentage vs UMI Count

```{r, eval=FALSE, fig.width=8, fig.height=6}
library(Seurat)
FeatureScatter(obj, feature1 = "percent.mt", feature2 = "nCount_RNA") +
  geom_smooth(method = "lm", color = "blue") +
  labs(x = "nCount_RNA", y = "percent.mt", 
         color = "blue") +
  theme_bw() +
  plot_annotation(title = "MT Percentage vs UMI Count")
```

## 3.2 Data Quality Metrics

### 3.2.1 Before vs After Comparison

| Metric | Before | After |
|--------|--------|-------|
| **Total Cells** | `r qc_stats[1, 3]` | `r qc_stats[1, 4]` |
| **Total Genes** | `r qc_stats[2, 4]` | `r qc_stats[2, 4]`` |
| **Mean Features/Cell** | `r qc_stats[3, 3]` | `r qc_stats[3, 3]`` |
| **Mean Counts/Cell** | `r qc_stats[5, 3]` | `r qc_stats[5, 3]` | `r qc_stats[6, 3]` |
| **Mean MT Percentage** | `r qc_stats[7, 3]` | `r qc_stats[8, 3]` | `r qc_stats[8, 3]`` |

# 4. Filtering Summary

## 4.1 Cell/Gene Filtering

**Low Quality Cells Removed:**

```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
library(Seurat)
# Assuming cells_removed is a vector of cell barcodes
cat("Low quality cells removed:", length(cells_removed), "\n")
```

## 5. Conclusions

Based on the QC metrics and filtering thresholds:

- Data quality is `r ifelse(qc_stats[8, 3] > config$qc$max_mt_pct, "above", "below")` the acceptable MT percentage threshold (`r config$qc$max_mt_pct`%`).
- `r ncol(qc_stats[1, 3])` cells passed quality control.
- `r qc_stats[1, 3] - qc_stats[1, 4]` cells were removed due to low quality.

## 6. Recommendations

1. Check data quality - If MT percentage is very high, consider investigating potential sample issues.
2. Adjust QC thresholds based on your data characteristics.
3. Review violin plots to identify outliers and consider adjusting filtering parameters.
4. Consider additional filtering steps if needed.

---

# Appendix: QC Parameters

| Parameter | Value |
|-----------|-------|--------|
| **Min Cells/Genes** | `r config$qc$min_cells_per_gene` |
| **Min Features/Cell** | `r config$qc$min_features_per_cell` |
| **Max Features/Cell** | `r ifelse(is.null(config$qc$max_features_per_cell), "Not set", config$qc$max_features_per_cell)` |
| **Max MT Percentage** | `r config$qc$max_mt_pct` |
| **Doublet Detection** | `r ifelse(config$qc$doublet_detection, "Enabled", "Disabled")` |

## Session Info

```{r, eval=FALSE, echo=FALSE}
sessionInfo()
```
