---
title: "Clustering Report"
author: "r-scrna development team"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

# Clustering Report

## Overview

This report summarizes the clustering analysis performed at multiple resolutions to identify optimal cell clusters.

## Analysis Information

- **Analysis Date**: `r format(Sys.Date(), "%Y-%m-%d")`
- **Pipeline Version**: r-scrna v1.0

## Clustering Parameters

| Parameter | Value | Description |
|-----------|--------|-------------|
| Resolutions Tested | `r params$clustering$resolutions` | Clustering resolutions evaluated |
| Algorithm | `r params$clustering$algorithm` | Clustering algorithm used |
| PC Dimensions | `r pcs_to_use` | Number of principal components used |
| Reduction | `r reduction` | Dimensionality reduction method |

## Clustering Metrics

```{r, eval=FALSE, echo=FALSE}
library(Seurat)
library(dplyr)
library(patchwork)
library(knitr)

# Assuming cluster_metrics and params are available from previous analysis
# This would be populated by the actual pipeline run
```

### Resolution-wise Cluster Counts

| Resolution | Number of Clusters | Mean Cells/Cluster | Std Cells/Cluster |
|-----------|---------------------|------------------|------------------|
`r table(cluster_metrics$resolution, cluster_metrics$n_clusters) %>% 
  kable(format = "markdown")```

### Overall Cluster Distribution

```{r, eval=FALSE, echo=FALSE}
library(Seurat)
library(dplyr)
library(ggplot2)

# Assuming cluster_summary from annotation module is available
cluster_props <- calc_cluster_props(obj, "cell_type")

ggplot(cluster_props, aes(x = cell_type, y = proportion, fill = cell_type)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_bw() +
  theme(axis.text = element_text(size = 12),
  legend.position = "bottom") +
  labs(title = "Cell Type Proportions", 
         x = "", 
         y = "Proportion")
)
```

### Clustering Quality Assessment

```{r, eval=FALSE, echo=FALSE}
library(Seurat)
library(patchwork)

# Assuming cluster_metrics from clustering module is available
# This would show metrics like silhouette scores, clumpiness scores, etc.
```

## Clustering Recommendations

Based on the analysis, the following clustering resolution is recommended:

- **Optimal Resolution**: `r params$clustering$resolutions[which.max(cluster_metrics$n_clusters[which.max(cluster_metrics$silhouette_score)))]`
- **Alternative**: `r params$clustering$resolutions[2]` (Second best resolution)

## Cluster Comparison Across Resolutions

### Pairwise Jaccard Similarity

```{r, eval=FALSE, echo=FALSE}
library(dplyr)

# Assuming multiple clustering results at different resolutions
# Calculate Jaccard similarity between clusterings

# This would show how stable clusters are across different resolutions
```

## Visualization: UMAP at Different Resolutions

```{r, eval=FALSE, echo=FALSE}
library(Seurat)
library(patchwork)

# Assuming we have Seurat object with multiple cluster assignments
for (res in params$clustering$resolutions) {
  cluster_col <- paste0("RNA_snn_res.", res)
  obj <- SetIdent(obj, value = cluster_col)
  
  p <- DimPlot(obj, reduction = "umap", group.by = cluster_col, label = TRUE, 
            label.size = 6, pt.size = 0.5, cols = get_colors()) +
    labs(title = paste0("UMAP at Resolution ", res), 
         subtitle = "Cells: ", ncol(obj))) +
    print(p)
}
```

## Resolution Selection Guidance

- **High Resolution** (e.g., 1.0): More fine-grained clusters, suitable for detecting subtle subpopulations
- **Medium Resolution** (e.g., 0.4-0.6): Balanced between granularity and stability
- **Low Resolution** (e.g., 0.8): More coarse clusters, suitable for major cell types

## Session Info

```{r, echo=FALSE}
sessionInfo()
```
