---
title: "Annotation Report"
author: "r-scrna development team"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

# Annotation Report

## Overview

This report summarizes cell type annotation results, including marker genes and cell type proportions.

## Analysis Information

- **Analysis Date**: `r format(Sys.Date(), "%Y-%m-%d")`
- **Pipeline Version**: r-scrna v1.0
- **Species**: `r config$species$name`
- **Genome**: `r config$species$genome`
- **Annotation Method**: `r config$annotation$method`

## Cell Type Statistics

### Cluster Summary Table

`r kable(cluster_summary)`

### Cell Type Proportions

```{r, echo=FALSE, fig.width=10, fig.height=6}
ggplot(cluster_props, aes(x = cell_type, y = proportion, fill = cell_type)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_bw() +
  theme(axis.text = element_text(size = 12)) +
  labs(x = "", y = "Proportion") +
  plot_annotation(title = "Cell Type Proportions")
```

## Marker Genes

### All Cluster Markers

```{r, echo=FALSE}
top_n_markers <- 10

# Extract top markers for each cluster
top_markers_list <- list()
for (cluster in unique(markers$cluster)) {
  cluster_markers <- markers %>%
    dplyr::filter(cluster == !!cluster) %>%
      dplyr::arrange(desc = avg_log2FC) %>%
      dplyr::slice_head(n = top_n_markers)
  top_markers_list[[cluster]] <- cluster_markers$gene
}

# Combine into data frame
all_top_markers <- dplyr::rbind_list(top_markers_list)

# Display as table
if (nrow(all_top_markers) > 0) {
  kable(all_top_markers)
}
```

### Cell Type-Specific Markers

```{r, echo=FALSE}
# Get markers for each cell type
cell_type_markers <- list()
for (cell_type in unique(markers$cell_type)) {
  type_markers <- markers %>%
    dplyr::filter(cell_type == !!cell_type) %>%
      dplyr::arrange(desc = avg_log2FC) %>%
      dplyr::slice_head(n = top_n_markers)
  cell_type_markers[[cell_type]] <- type_markers$gene
}

# Display as table by cell type
if (length(cell_type_markers) > 0) {
  cat("Cell Type-Specific Markers:\n\n")
  for (ct in names(cell_type_markers)) {
    cat("\n", ct, ":\n")
    print(kable(head(cell_type_markers[[ct]], n = 10))
  }
}
```

## Visualization

### UMAP with Annotations

```{r, echo=FALSE, fig.width=10, fig.height=8}
p1 <- DimPlot(obj, reduction = "umap", group.by = "cell_type", 
             label = TRUE, label.size = 6, pt.size = 0.5) +
             cols = get_colors())

p2 <- DimPlot(obj, reduction = "umap", group.by = "seurat_clusters", 
             label = TRUE, label.size = 6, pt.size = 0.5, 
             cols = get_colors())

p3 <- FeaturePlot(obj, features = top10_genes, 
                  group.by = "cell_type", pt.size = 0.3) +
             cols = get_colors())

combined <- (p1 + p2) / p3
print(combined)
```

## Session Info

```{r, echo=FALSE}
sessionInfo()
```
