---
title: "Differential Expression Report"
author: "r-scrna development team"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

# Executive Summary

## Analysis Overview

Differential expression analysis was performed to identify genes that are differentially expressed between groups.

## Analysis Parameters

| Parameter | Value |
|-----------|--------|
| **Test Used** | `r params$de$test_use` |
| **Minimum P-value** | `r params$de$pvalue_cutoff` |
| **Log2FC Threshold** | `r params$de$logfc_threshold` |
| **Minimum PCT** | `r params$de$min_pct` |
| **Only Positive** | `r params$de$only_pos` |

## Comparison Summary

### Number of Comparisons Performed: `r length(de_results)`

### Groups Analyzed: `r paste(groups_analyzed, collapse = ", ")`

### Up-regulated Genes

| Cluster | Down-regulated Genes |
|---------|-------------------|
`r sapply(de_results, function(x) { sum(x$avg_log2FC > 0) })` | `r sapply(de_results, function(x) { sum(x$avg_log2FC < 0) })` `

### Top DEG Count by Cluster

| Cluster | Total DEGs | Up | Down |
|---------|----------|------|
`r bind_rows(de_results, de_results %>% 
      dplyr::group_by(cluster) %>%
      dplyr::summarise(Total_DEGs = n(), Up = n_up, Down = n_down, .groups = "cluster")) %>%
      arrange(desc(Total_DEGs)) %>%
      kable(caption = "Top DEG Count by Cluster")
  `

---

# Visualization: Volcano Plot

## Volcano Plot

The volcano plot shows significance (adjusted p-value) vs log2 fold change for each comparison.

## Top Table: Top Differentially Expressed Genes

```{r}
top20_all <- de_results %>% 
  dplyr::arrange(desc = abs(avg_log2FC)) %>%
  slice_head(n = 20) %>%
  dplyr::mutate(
    Regulation = dplyr::case_when(
      avg_log2FC > 0 ~ "Up",
      avg_log2FC < 0 ~ "Down"
    ),
    significance = dplyr::case_when(
      p_val_adj < params$de$pvalue_cutoff ~ "Significant",
      TRUE ~ "NS"
    )
  )

kable(top20_all, caption = "Top 20 DEGs (All Comparisons Combined)", 
      digits = 4,
      format = "markdown")
```

## Top Table: Top Up-regulated Genes

```{r}
top20_up <- de_results %>% 
  dplyr::filter(avg_log2FC > 0) %>%
  dplyr::arrange(desc = desc(avg_log2FC)) %>%
  slice_head(n = 20)

kable(top20_up, caption = "Top 20 Up-regulated Genes", 
      digits = 4,
      format = "markdown")
```

## Top Table: Top Down-regulated Genes

```{r}
top20_down <- de_results %>% 
  dplyr::filter(avg_log2FC < 0) %>%
  dpr::arrange(desc = desc(avg_log2FC)) %>%
  slice_head(n = 20)

kable(top20_down, caption = "Top 20 Down-regulated Genes", 
      digits = 4,
      format = "markdown")
```

---

# MA Plot

## MA Plot

The MA plot shows the relationship between mean expression and fold change.

## Heatmap: Top DEG Heatmap

Heatmap showing expression of top DEGs.

```{r, fig.width=12, fig.height=8, fig.align='center'}
if (exists("de_heatmap_data") {
  de_heatmap_data <- de_heatmap_data
} else {
  # Get top 50 DEGs
  all_degs <- de_results %>% dplyr::arrange(desc = abs(avg_log2FC)) %>% slice_head(n = 50)
  
  # Get expression data
  expr_data <- GetAssayData(obj, slot = "scale.data")[rev(all_degs$gene), ]
  
  # Create annotation
  anno_df <- data.frame(
    Cluster = all_degs$cluster,
    Regulation = dplyr::case_when(
      avg_log2FC > 0 ~ "Up",
      avg_log2FC < 0 ~ "Down",
      TRUE ~ "NS"
    )
  )
  
  # Plot heatmap
  pheatmap(
    as.matrix(expr_data),
    cluster_rows = FALSE,
    cluster_cols = FALSE,
    show_colnames = FALSE,
    show_rownames = FALSE,
    annotation_row = anno_row,
    cluster_cols = c("Up" = "#E41A36", "Down" = "#3366CC"),
    cluster_col = "Cluster",
    border_color = "white",
    fontsize = 8,
    show_colnames = FALSE,
    scale = "row"
  )
}
```

## Cluster Comparison Heatmap

Heatmap comparing DEGs between clusters.

---

# Enrichment Analysis

## GO Enrichment

### Summary Statistics

| Metric | Value |
|--------|-------|
| **Total GO Terms Found** | `r nrow(ego@result)` |
| **Significant Terms (adj.p < `params$enrichment$pvalue_cutoff`)** | `r sum(ego@result$p.adjust < params$enrichment$pvalue_cutoff)` |
| **Up-regulated GO Terms** | `r sum(ego@result$p.adjust < params$enrichment$pvalue_cutoff & ego@result$logFC > 0)` |
| **Down-regulated GO Terms** | `r sum(ego@result$p.adjust < params$enrichment$pvalue_cutoff & ego@result$logFC < 0)` |

### GO Enrichment Results

```{r}
kable(ego@result[1:min(20, nrow(ego@result)), 
      caption = "GO Enrichment Results (Top 20)",
      digits = 4,
      format = "markdown")
  `

# GO Plot
go_bar <- create_go_barplot(ego, n_terms = 20, config)
print(go_bar)
```

## KEGG Enrichment

### Summary Statistics

| Metric | Value |
|--------|-------|
| **Total KEGG Pathways Found** | `r nrow(kegg@result)` |
| **Significant Pathways (adj.p < `params$enrichment$pvalue_cutoff`)** | `r sum(kegg@result$p.adjust < params$enrichment$pvalue_cutoff)` |
| **Up-regulated Pathways** | `r sum(kegg@result$p.adjust < params$enrichment$pvalue_cutoff & kegg@result$logFC > 0)` |
| **Down-regulated Pathways** | `r sum(kegg@result$p.adjust < params$enrichment$pvalue_cutoff & kegg@result$logFC < 0)` |

### KEGG Enrichment Results

```{r}
kable(kegg@result[1:min(20, nrow(kegg@result)), 
      caption = "KEGG Enrichment Results (Top 20)",
      digits = 4,
      format = "markdown"
  `

# KEGG Plot
kegg_bar <- create_kegg_barplot(kegg, n_terms = 20, config)
print(kegg_bar)
```

## Enrichment Comparison

```{r, eval=FALSE}
if (exists("go_enrichment_comparison")) {
  print(go_enrichment_comparison)
}
```

---

# Session Info

```{r}
sessionInfo()
```
