---
title: "r-scrna Full Analysis Report"
author: "r-scrna development team"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

# Executive Summary

This report presents a comprehensive analysis of single-cell RNA-seq data using the **r-scrna** automated pipeline.

## Project Information

- **Analysis Date**: `r format(Sys.Date(), "%Y-%m-%d")`
- **Pipeline Version**: r-scrna v1.0
- **Species**: `r config$species$name`
- **Reference Genome**: `r config$species$genome`
- **Data Type**: `r config$input$data_type`

## Analysis Workflow

```
00. Input Checks  → 01. Quality Control  → 02. Normalization  → 03. Clustering → 04. Annotation → 05. Differential Expression → 06. Enrichment → [07. Pseudotime] → [08. CellChat] → [09. SCENIC] → [10. RNA Velocity] → [11. CNV] → 09. Report
```

## Table of Contents

1. [Data Quality](#data-qc-report)
2. [Clustering](#clustering-report)
3. [Annotation](#annotation-report)
4. [Differential Expression](#diffexp-report)
5. [Enrichment](#enrichment-report)
6. [Optional Analyses](#optional-report)

---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(rmarkdown)
library(patchwork)

# Read configuration
config <- yaml::read_yaml("config/config.yaml")

# Check which modules were run
module_run_status <- list(
  data_qc = file.exists("outputs/inter/qc_filtered.rds"),
  normalization = file.exists("outputs/inter/normalized.rds"),
  clustering = file.exists("outputs/inter/clustered.rds"),
  annotation = file.exists("outputs/inter/annotated.rds"),
  diffexp = file.exists("outputs/enrichment/de_results.csv"),
  enrichment = file.exists("outputs/enrichment/enrichment_results.csv"),
  pseudotime = !is.null(config$modules$pseudotime) || !config$modules$pseudotime$enabled,
  cellchat = !is.null(config$modules$cellchat) || !config$modules$cellchat$enabled,
  scenic = !is.null(config$modules$scenic) || !config$modules$scenic$enabled,
  rna_velocity = !is.null(config$modules$rna_velocity) || !config$modules$rna_velocity$enabled,
  cnv = !is.null(config$modules$cnv) || !config$modules$cnv$enabled
)
)

# Include optional modules only if they were run
include_optional <- pseudotime || cellchat || scenic || rna_velocity || cnv
```

---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
# Read intermediate objects if they exist
qc_obj <- if (module_run_status$data_qc) readRDS("outputs/inter/qc_filtered.rds")
norm_obj <- if (module_run_status$normalization) readRDS("outputs/inter/normalized.rds")
cluster_obj <- if (module_run_status$clustering) readRDS("outputs/inter/clustered.rds")
annotated_obj <- if (module_run_status$annotation) readRDS("outputs/inter/annotated.rds")

# Read analysis results if available
de_results <- NULL
if (module_run_status$diffexp) {
  de_results <- read.csv("outputs/enrichment/de_results.csv", row.names = 1)
}

enrichment_results <- NULL
if (module_run_status$enrichment) {
  enrichment_results <- read.csv("outputs/enrichment/enrichment_results.csv", row.names = 1)
}
```

# Module 1: Data Quality Control
```{r data-qc-report, child='results="data-qc-report", include=if(module_run_status$data_qc)}
```

`r if (module_run_status$data_qc) "Read intermediate QC object" else "QC module was not run"`

### Initial Data Statistics

| Metric | Value |
|--------|--------|
`r if (module_run_status$data_qc) {
  qc_stats <- readRDS("outputs/inter/qc_filtered.rds")@metadata$qc_stats
} else {
  qc_stats <- data.frame(
    metric = c("Total Cells", "Total Genes", "Mean Features/Cell", "Median Features/Cell", 
             "Mean Counts/Cell", "Median Counts/Cell", 
             "Mean MT Percentage", "Median MT Percentage"),
    value = c(
      NA, NA, NA, NA, NA, NA, NA, NA
    )
  )
}
print(knitr::kable(qc_stats))```

### Filtering Results

| Parameter | Threshold | Cells Filtered |
|-----------|----------|-------------|
`r if (module_run_status$data_qc) {
  config_str <- sprintf("Min Features/Cell: %d, Max Features/Cell: %d, Max MT PCT: %.1f%%", 
                                 config$qc$min_features_per_cell, 
                                 config$qc$max_features_per_cell, 
                                 config$qc$max_mt_pct * 100)
  print(sprintf(config_str, "", knitr::knit))
}`

### QC Visualizations

```{r data-qc-report, child='results="data-qc-report", include=if(module_run_status$data_qc)}
```
`r if (module_run_status$data_qc) {
  qc_plots <- readRDS("outputs/inter/qc_filtered.rds")@metadata$qc_plots
  combined_qc_plots <- qc_plots$combined
  print(combined_qc_plots)
}`

---

# Module 2: Clustering
```{r clustering-report, child='results="clustering-report", include=if(module_run_status$clustering)}
```

`r if (module_run_status$clustering) "Read intermediate clustering object" else "Clustering module was not run"`

### Clustering Metrics

| Resolution | N Clusters | Mean Silhouette Score |
|-----------|-----------|-------------|----------------|
`r if (module_run_status$clustering) {
  cluster_metrics <- readRDS("outputs/inter/clustered.rds")@metadata$cluster_metrics
  print(knitr::kable(cluster_metrics))
}````

### UMAP Visualization

```{r clustering-report, child='results="clustering-report", include=if(module_run_status$clustering)}
```
`r if (module_run_status$clustering) {
  umap_plots <- readRDS("outputs/inter/clustered.rds")@metadata$umap_plots
  combined_umap <- readRDS("outputs/inter/clustered.rds")@metadata$combined_umap)
  print(combined_umap)
}````

---

# Module 3: Annotation
```{r annotation-report, child='results="annotation-report", include=if(module_run_status$annotation)}
```

`r if (module_run_status$annotation) "Read intermediate annotated object" else "Annotation module was not run"`

### Cell Type Summary

```{r annotation-report, results="annotation-report", include=if(module_run_status$annotation)}
`r if (module_run_status$annotation) {
  cluster_summary <- readRDS("outputs/inter/annotated.rds")@metadata$cluster_summary
  print(knitr::kable(cluster_summary))
```

### Marker Genes Summary

```{r annotation-report, results="annotation-report", include=if(module_run_status$annotation)}
`r if (module_run_status$annotation) {
  markers <- readRDS("outputs/inter/annotated.rds")@metadata$markers
  top5_per_cluster <- markers %>%
    dplyr::group_by(cluster) %>%
    dplyr::slice_head(n = 5)
  
  p <- DoHeatmap(obj, features = top5$gene, 
             size = 3, group.by = "cell_type",
             angle = 90, pt.size = 0.5) +
    plot_annotation(title = "Top 5 Markers per Cluster")
  print(p)
}````

---

# Module 4: Differential Expression
```{r diffexp-report, child='results="diffexp-report", include=if(module_run_status$diffexp)}
```

`r if (module_run_status$diffexp) "Read DE results" else "DEG module was not run"`

### DEG Summary

| Comparison | Up-regulated Genes | Down-regulated Genes |
|-----------|-------------------|-------------------|----------------|
`r if (module_run_status$diffexp) {
  total_up <- sum(de_results$regulated == "UP")
  total_down <- sum(de_results$regulated == "Down")
  
  df_summary <- data.frame(
    comparison = de_results$comparison,
    n_up = sapply(de_results, function(x) nrow(subset(x, regulated == "UP")),
    n_down = sapply(de_results, function(x) nrow(subset(x, regulated == "Down"))
  )
  print(knitr::kable(df_summary))
}````

### Top Markers by Cluster

```{r diffexp-report, results="diffexp-report", include=if(module_run_status$diffexp)}
`r if (module_run_status$diffexp) {
  markers <- readRDS("outputs/inter/annotated.rds")@metadata$markers
  top5_per_cluster <- markers %>%
    dplyr::group_by(cluster) %>%
    dplyr::slice_head(n = 5) %>%
    dplyr::select(cluster, gene, avg_log2FC)
  print(knitr::kable(top5_per_cluster))
}````

### Volcano Plot

```{r diffexp-report, results="diffexp-report", include=if(module_run_status$diffexp)}
```
`r if (module_run_status$diffexp) {
  volcano_p <- readRDS("outputs/enrichment/de_results.rds")@metadata$plots$volcano
  print(volcano_p)
}````

### Heatmap of Top Genes

```{r diffexp-report, results="diffexp-report", include=if(module_run_status$diffexp)}
`r if (module_run_status$diffexp) {
  de_heatmap <- readRDS("outputs/enrichment/de_results.rds")@metadata$plots$heatmap
  print(de_heatmap)
}````

---

# Module 5: Enrichment
```{r enrichment-report, child='results="enrichment-report", include=if(module_run_status$enrichment)}
```

`r if (module_run_status$enrichment) "Read enrichment results" else "Enrichment module was not run"`

### GO Enrichment Summary

| Database | Species | P-value Cutoff | Q-value Cutoff | Min Geneset Size | Pathways Enriched |
|-----------|----------|----------|----------|-------|-----------------------|---------------------|-----------|
`r if (module_run_status$enrichment) {
  go_summary <- read.csv("outputs/enrichment/go_summary.csv", row.names = 1)
  print(knitr::kable(go_summary))
}````

### KEGG Enrichment Summary

| Database | P-value Cutoff | Q-value Cutoff | Min Geneset Size | Pathways Enriched |
|-----------|----------|----------|----------|-------|---------------------|-----------|
`r if (module_run_status$enrichment) {
  kegg_summary <- read.csv("outputs/enrichment/kegg_summary.csv", row.names = 1)
  print(knitr::kable(kegg_summary))
}````

### Enrichment Visualization

```{r enrichment-report, results="enrichment-report", include=if(module_run_status$enrichment)}
`r if (module_run_status$enrichment) {
  enrich_plots <- readRDS("outputs/enrichment/enrichment/enrichment_plots")
  go_bar <- enrich_plots$go_bar
  kegg_bar <- enrich_plots$kegg_bar
  go_dot <- enrich_plots$go_dot
  kegg_dot <- enrich_plots$kegg_dot
  comparison_p <- enrich_plots$comparison
  
  combined_enrichment <- wrap_plots(list(go_bar, kegg_bar, go_dot, kegg_dot))
  print(combined_enrichment)
}````

---

# Optional Analyses

```{r optional-report, child='results="optional-report", include=if(include_optional)}
```

## Module 7: Pseudotime Analysis

`r if (pseudotime) "### 7. Pseudotime Analysis\n" else ""`

```{r optional-report, child='results="optional-report", include=if(pseudotime)}
```
`r if (pseudotime) {
  pseudotime_obj <- readRDS("outputs/inter/pseudotime/pseudotime.rds")
  print(head(pseudotime_obj))
}`

---

## Module 8: CellChat Analysis

`r if (cellchat) "### 8. CellChat Analysis\n" else ""`

```{r optional-report, child='results="optional-report", include=if(cellchat)}
```
`r if (cellchat) {
  cellchat_obj <- readRDS("outputs/inter/cellchat/cellchat_obj")
  print(head(cellchat_obj))
}````

---

## Module 9: SCENIC Analysis

`r if (scenic) "### 9. SCENIC Analysis\n" else ""`

```{r optional-report, child='results="optional-report', include=if(senic)}
```
`r if (scenic) {
  scenic_obj <- readRDS("outputs/inter/scenic/scenic_obj")
  print(head(scenic_obj))
}````

---

## Module 10: RNA Velocity Analysis

`r if (rna_velocity) "### 10. RNA Velocity Analysis\n" else ""`

```{r optional-report, child='results="optional-report', include=if(rna_velocity)}
```
`r if (rna_velocity) {
  velocity_obj <- readRDS("outputs/inter/rna_velocity/rna_velocity_obj")
  print(head(velocity_obj))
}````

---

# Module 11: CNV Analysis

`r if (cnv) "### 11. CNV Analysis\n" else ""`

```{r optional-report, child='results="optional-report', include=if(cnv)}
```
`r if (cnv) {
  cnv_obj <- readRDS("outputs/inter/cnv/cnv_obj")
  print(head(cnv_obj))
}````

---

# Session Info

```{r optional-report, results="optional-report"}
```{r}
sessionInfo()
```
